\thetitle

### 2008-06-05 paper.tex
Overhauling Related Work: Harpoon.

The second reason why uniformity and marginality assumptions are so commonly used is that there are no generally accepted alternatives for representing or analyzing traffic.
The difficulty lies in the high-dimensional and stochastic nature of network behaviors.
Suppose, for example, that we assume that packet sizes and inter-packet intervals are drawn independently from some given distribution for each flow.
In the terminology introduced above, this is a flow-conditional packet behavior model.
Prior research has indicated that, given accurate per-flow distributions, this is a sufficiently realistic model of network behavior~\cite{Karpinski07:realism}.\realismnote
In this model, however, each flow must have its own specific, realistic pair of distributions for packet size and inter-packet intervals.
Moreover, the ``distribution of distributions'' across the entire network must simultaneously be realistic.
Modeling such distributions of distributions realistically for entire networks is currently an unsolved problem; we provide the means to address it.
% This paper provides the mathematical framework in which to address this problem and demonstrates the effectiveness of this framework by taking the first steps towards fully understanding the complex interdependencies of real network behavior.
% make a major first step towards being able to model the distribution of flow behaviors across an entire network, by providing the framework in which to address this problem mathematically, and by demonstrating that this framework actually allows us to meaningfully analyze non-uniform, non-independent cross-network behavior.

@@

We argue that matrix factorization, which is nondestructive, is a far superior to multiplication as a way to reduce the dimensionality of network traffic behaviors.
% To show how traffic can be modeled less destructively, 
% To demonstrate how linear representation allows drastically improved analysis and modeling of network traffic,
To demonstrate this approach, we use non-negative matrix factorization (\caps{NMF}) to find a low-dimensional approximation of the packet size and inter-packet interval behavior matrices.
% As an example of how the linear representation allows drastically improved analysis and modeling of network traffic, we apply non-negative matrix factorization to find an low-dimensional representation of packet size and inter-packet flow behaviors.
% Each flow's behavior can be explained as a linear combination of only twelve ``basic behaviors.'' 
% Although they are found purely algorithmically from packet size and inter-packet interval distributions, these behaviors correspond to intuitive mode of network activity.
This factorization extracts a small set of ``basic behaviors'' from a traffic matrix and simultaneously explains each flow's observed behavior as a mixture of these basic behaviors.
Although the basic behaviors are found without examining port numbers or protocols, they nevertheless correspond to intuitive network activities such as file transfer, typing via \caps{SSH}, web surfing, and ping traffic.

@@

Paxson and Floyd observed~\cite{Paxson95} that the interplay between endpoint behavior and the network conditions is inherently \textit{closed-loop} in the sense that it is potentially affected by complex feedback. Traffic models typically attempt to preserve the closed-loop behavior of network traffic. Characterizing and modeling the feedback between the network and applications using \caps{TCP} is precisely the subject of Hern\'andez-Campos's dissertation~\cite{Hernandez06:dissertation}. Hern\'andez-Campos's addresses in great depth how to abstract an application's behavior from a \caps{TCP} trace and replay it in a manner that accurately reflects how most common applications \emph{would} react in response to other network conditions. This research has subsequently been turned into a \caps{TCP} flow generation tool for \caps{NS-2}, called \tmix~\cite{Weigle06}. It is important to note that {\tmix} does not provide models for application behavior or for mixtures of applications in networks. Rather, it provides the ability to reproduce accurate closed-loop dynamics for a single \caps{TCP} flow \emph{given} an application behavior model as input. Our work complements this perfectly: it provides the means to analyze traffic and generate realistic multi-flow application workloads which can then serve as the necessary input to \tmix.

@@

There is a large and diverse body of work on traffic analysis, modeling, and generation~\cite{Paxson95,Paxson96,Sommers04,Avallone04,Hernandez06:dissertation}. We are only able to discuss a small, but hopefully representative sampling of this work. Almost all of the traffic generation work has focused on wide-area Internet backbone traffic.
%Because of the comparatively unlimited capacity of wired networks, traffic workload generally does not offer interesting technical challenges at the edge of the network.
The two most prominent traffic generation frameworks are Harpoon and \caps{D-ITG}. Harpoon~\cite{Sommers04} uses a traffic trace for self-training, and can subsequently generate synthetic traffic with certain statistical properties based on the original trace. The properties reproduced are the empirical distributions of the following: ``file size, inter-connection time, source and destination IP ranges, number of active sessions.'' There is no criterion proposed to determine whether these properties characterize the original traffic adequately---we can only hope that this approximation is good enough. For many purposes, it likely is sufficient; in particular, it is probably appropriate for the intended use in generating traffic for Internet backbone simulations. Wireless networks, however, are particularly sensitive to workload conditions, and sampling from a limited set of empirical distributions does not suffice to reproduce realistic network-wide traffic.

@@

Finally, we turn to the regularized models. These demonstrate neither exceptionally good nor exceptionally bad behavior. That the regularized models should be similar to the \model{full marginal} model but somewhat more realistic is entirely expected: regularization, the reader will recall, is a restricted form of marginal modeling, where each group of flows has its own marginal distribution. In these cases, the flows are grouped by source node or by destination node. The mixed case uses the average behavior of the other two cases. It is interesting to note that the mixed case is universally worse than either the source or destination case. Clearly, hybridizing the regularization here is not an effective technique. We believe that the cause of this effect is like mixing too many colors yielding an unpleasant shade of brown: any of the colors individually would be better than the non-descript average. As a more general conclusion, arbitrarily mixing model may not always be a good idea, despite the ease with which it is done in the matrix representation of workload behavior.

>>>>

The second reason why uniformity and marginality assumptions are so commonly used is that there are no generally accepted alternatives for representing or analyzing traffic.
The difficulty lies in the high-dimensional and stochastic nature of network behaviors.
Suppose, for example, that we assume that packet sizes and inter-packet intervals are drawn independently from some given distribution for each flow.
In the terminology introduced above, this is a flow-conditional packet behavior model.
Prior research has indicated that, given accurate per-flow distributions, this is a sufficiently realistic model of network behavior~\cite{Karpinski07:realism}.\realismnote
In this model, however, each flow must have its own specific, realistic pair of distributions for packet size and inter-packet intervals.
Moreover, the ``distribution of distributions'' across the entire network must simultaneously be realistic.
Modeling such distributions of distributions realistically for entire networks is currently an unsolved problem; we provide the means to address it.

@@

We argue that matrix factorization, which is nondestructive, is a far superior to multiplication as a way to reduce the dimensionality of network traffic behaviors.
To demonstrate this approach, we use non-negative matrix factorization (\caps{NMF}) to find a low-dimensional approximation of the packet size and inter-packet interval behavior matrices.
This factorization extracts a small set of ``basic behaviors'' from a traffic matrix and simultaneously explains each flow's observed behavior as a mixture of these basic behaviors.
Although the basic behaviors are found without examining port numbers or protocols, they nevertheless correspond to intuitive network activities such as file transfer, typing via \caps{SSH}, web surfing, and ping traffic.

@@

\subsection{Traffic Generation}

Paxson and Floyd observed~\cite{Paxson95} that the interaction between endpoint behavior and the network conditions is inherently \textit{closed-loop} in the sense that the exhibited behavior is governed by a closed feedback loop.
Traffic models typically attempt to preserve the closed-loop behavior of network traffic.
In the case of most non-\caps{TCP} applications, this feedback does not occur: an open-loop model suffices.
Hern\'andez-Campos's dissertation~\cite{Hernandez06:dissertation} addresses in great depth how to model the feedback between the network and applications for most \caps{TCP} applications.
Hern\'andez-Campos shows how to abstract an application's behavior from a \caps{TCP} trace and replay it in a manner that accurately reflects how most common applications would react in response to different network conditions.
This research has subsequently been turned into a \caps{TCP} flow generation tool called \tmix~\cite{Weigle06}.
It is important to note that {\tmix} does not provide models for application behaviors or for mixtures of applications in networks.
Rather, it provides the ability to reproduce accurate closed-loop dynamics for a single \caps{TCP} flow \emph{given} an application behavior model as input.
Our work complements this perfectly: this paper provides the means to analyze traffic and generate realistic whole-network application workloads which can then serve as the necessary input to \tmix.

@@

% There is a large and diverse body of work on traffic analysis, modeling, and generation~\cite{Paxson95,Paxson96,Sommers04,Avallone04,Hernandez06:dissertation}. We are only able to discuss a small, but hopefully representative sampling of this work. Almost all of the traffic generation work has focused on wide-area Internet backbone traffic.
%Because of the comparatively unlimited capacity of wired networks, traffic workload generally does not offer interesting technical challenges at the edge of the network.
The two most prominent general traffic generation frameworks are Harpoon and \caps{D-ITG}. Harpoon~\cite{Sommers04} uses a traffic trace for self-training, and can subsequently generate synthetic traffic with certain statistical properties based on the original trace.
The properties reproduced are the empirical distributions of the following: ``file size, inter-connection time, source and destination IP ranges, number of active sessions.''
Harpoon exemplifies the uniform and marginal modeling approaches.
Source and destination nodes are treated marginally: an empirically derived marginal frequencies of sources and destination nodes (independently) are used to choose endpoints for each flow.
File sizes (\latin{i.e.} total flow sizes) are likewise treated marginally: each flow's total byte  count is sampled from a marginal distribution of flow sizes.
\caps{UDP} flows are modeled as being constant bit-rate (\latin{i.e.} having uniform packet behavior), and \caps{TCP} flows are all treated as file transfers, making them effectively uniform with bandwidth limited by simulated \caps{TCP} dynamics.
Inter-connection times (\latin{i.e.} per node-pair inter-flow intervals) are sampled from a single empirically derived marginal distribution.
While Harpoon may be sufficiently realistic for the intended purpose of generating Internet backbone traffic, it is difficult to know this:
generated traffic is only compared with trace traffic using metrics that are explicitly specified in the models.
Realism is never double-checked using derived statistical metrics or actual performance metrics.
For the purpose of generating local-area traffic for wireless experiments, it is certainly not sufficiently realistic. Previous research~\cite{Karpinski07:cbr-failure} has shown that \caps{CBR} packet behavior models, like those used in Harpoon, drastically distort vital performance metrics at all levels of the protocol stack in wireless network simulations. This paper gives equal cause for concern about marginal traffic models.

% There is no criterion proposed to determine whether these properties characterize the original traffic adequately---we can only hope that this approximation is good enough. For many purposes, it likely is sufficient; in particular, it is probably appropriate for the intended use in generating traffic for Internet backbone simulations. Wireless networks, however, are particularly sensitive to workload conditions, and sampling from a limited set of empirical distributions does not suffice to reproduce realistic network-wide traffic.

@@

Finally, we turn to the regularized models. These demonstrate neither exceptionally good nor exceptionally bad behavior. That the regularized models should be similar to the \model{full marginal} model but somewhat more realistic is entirely expected: regularization, the reader will recall, is a restricted form of marginal modeling, where each group of flows has its own marginal distribution. In these cases, the flows are grouped by source node or by destination node. The mixed case uses the average behavior of the other two cases. It is interesting to note that the mixed case is universally worse than either the source or destination case. Clearly, hybridizing the regularization here is not an effective technique. We believe that the cause of this effect is like mixing too many colors yielding an unpleasant shade of brown: any of the colors individually would be better than the nondescript average. As a more general conclusion, arbitrarily mixing model may not always be a good idea, despite the ease with which it is done in the matrix representation of workload behavior.
